Drought vulnerability of tree growth

### 2024-07-08 scripts/explore_analysis.Rmd
updated script for explore plots

dendro_inc<- dendro.data %>%
  filter(Cno %in% alt_census)%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-lag(Cno),
         #removing increment value if the previous measure was before the last year
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)
```
```{r species wise increments, echo=F}

@@

saccli_plot<-ggplot() +
  geom_line(data = saccli, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag)) +
  geom_line(data=saccli %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for SACCLI (deciduous)")+
  theme_bw()

hopeod_plot<-ggplot() +
  geom_line(data = hopeod, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag)) +
  geom_line(data=hopeod %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for HOPEOD (evergreen)")+
  theme_bw()

tetrnu_plot<-ggplot() +
  geom_line(data = tetrnu, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag)) +
  geom_line(data=tetrnu %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for TETRNU (deciduous)")+
  theme_bw()

polyvi_plot<-ggplot() +
  geom_line(data = polyvi, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag)) +
  geom_line(data=polyvi %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for POLYVI (evergreen)")+
  theme_bw()

png("growth-precip-thailand/doc/display/ind_increments.png", width=8, height=8, units="in", res=300)
grid.arrange(hopeod_plot, tetrnu_plot, polyvi_plot, saccli_plot, nrow=2)
dev.off()

@@

@@

growth_top4
```

@@

png("growth-precip-thailand/doc/display/resistance_map_top4.png", width=8, height=8, units="in", res=300)
resistance_map
dev.off()

>>>>

dendro_inc<- dendro.data %>%
  filter(Cno %in% alt_census)%>%
  arrange(Cno)%>%
  group_by(Tag) %>% #grouping by treeID to calculate increment
  dplyr::mutate(calcDBH_min1=dplyr::lag(calcDBH, n=1, default=NA, order_by=Tag),
         cii_min1=dplyr::lag(Cii, n=1, default=NA, order_by=Tag),
         tdif=difftime(Date2, dplyr::lag(Date2, n=1, default=NA, order_by=Tag), units = "days"),
         inc=calcDBH-calcDBH_min1,
         inc_annual=(inc/as.numeric(tdif))*365,
         inc.time=Cno-dplyr::lag(Cno, n=1, default=NA, order_by=Tag),
         #removing increment value if the previous measure was before the last year
         inc_annual=ifelse(inc.time==2, inc_annual, NA),
         treeID=paste0("HKK.", Tag))%>%
  filter(!is.na(inc))%>%
  filter(!is.na(inc_annual))%>%
  ungroup()%>%
  dplyr::mutate(
    diainc.scaled = scale(inc_annual),
    out.inc = ifelse(abs(diainc.scaled)>3, 1, 0))%>%
  filter(out.inc==0)
```
```{r species wise increments, echo=F}

@@

saccli_plot<-ggplot() +
  geom_line(data = saccli, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=saccli %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=saccli %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for SACCLI (deciduous)")+
  theme_bw()

hopeod_plot<-ggplot() +
  geom_line(data = hopeod, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=hopeod %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=hopeod %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for HOPEOD (evergreen)")+
  theme_bw()

tetrnu_plot<-ggplot() +
  geom_line(data = tetrnu, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=tetrnu %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=tetrnu %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for TETRNU (deciduous)")+
  theme_bw()

polyvi_plot<-ggplot() +
  geom_line(data = polyvi, aes(x=((Cno-3)/2)+2009, y=inc_annual, group=Tag), alpha=0.5) +
  geom_line(data=polyvi %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=mean_inc), col="red")+
  geom_line(data=polyvi %>% group_by(Cno) %>% dplyr::summarise(median_inc=median(inc_annual, na.rm=T)), aes(x=((Cno-3)/2)+2009, y=median_inc), col="blue")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for POLYVI (evergreen)")+
  theme_bw()

library(gridExtra)
png("growth-precip-thailand/doc/display/ind_increments.png", width=8, height=8, units="in", res=300)
grid.arrange(hopeod_plot, tetrnu_plot, polyvi_plot, saccli_plot, nrow=2)
dev.off()

@@

#spaghetti plot of all species together

spagplot_top4<-ggplot() +
  #species plots
  geom_line(data=top4 %>% group_by(Cno, sp) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc, 
  group=factor(sp,levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI")), 
  col=factor(sp,levels=c("HOPEOD", "TETRNU", "POLYVI", "SACCLI"))))+
  geom_line(data=top4 %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="black", size=2)+
  scale_color_viridis_d()+
  guides(col=guide_legend(title="species"))+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 4 species")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_top4.png", width=4, height=4, units="in", res=300)
spagplot_top4
dev.off()

spagplot_all<-ggplot() +
  #species plots
  geom_line(data=dendro_inc %>% group_by(Cno, sp) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc, 
  group=sp, col=sp))+
  #mean of all trees
  geom_line(data=dendro_inc %>% group_by(Cno) %>% dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="black", size=2)+
  #mean of species
  geom_line(data=dendro_inc %>% group_by(Cno, sp) %>% 
  dplyr::summarise(mean_inc=mean(inc_annual, na.rm=T)) %>%
  ungroup()%>%
  group_by(Cno) %>% dplyr::summarise(mean_inc=mean(mean_inc, na.rm=T)), 
  aes(x=((Cno-3)/2)+2009, y=mean_inc), col="grey40", size=0.8)+
  scale_color_viridis_d()+
  guides(col="none", group="none")+
  xlab("year") + ylab("diameter increment (cm)")+
  ggtitle("growth increments for top 4 species")+
  theme_bw()

png("growth-precip-thailand/doc/display/spaghetti_all.png", width=4, height=4, units="in", res=300)
spagplot_all
dev.off()

@@

@@

png("growth-precip-thailand/doc/display/resistance_map_top4.png", width=10, height=8, units="in", res=300)
resistance_map
dev.off()
